---
title: "Extraction of M and SDs from Coles et al. 2024 for demand characteristics meta-analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---
In this study, participants completed a line judgment task.

The study featured a 2 (no information received; information received) x 2 (information received is about the researcher's hypothesis; information received is about typical group responses) x 2 (lines are left justified first; lines are right justified first) fractional factorial design. The first factor was manipulated within-subjects, and the second and third were manipulated between-subjects.

The condition where participants receive information about the researcher's hypothesis (vs. the condition where they receive no information) meets the inclusion criteria for the meta-analysis.

As of June 4, 2024, data for this study have been collected at 6 sites.

```{r}
# set up packages
library(tidyverse)
library(qualtRics)
```
 
# Open and clean data
```{r}
# Dangs API credential at UCSD (hidden for sharing)
qualtrics_api_credentials(
  api_key = 'HIDDEN', 
  base_url = 'HIDDEN',
  install = TRUE,
  overwrite = TRUE)

readRenviron("~/.Renviron")

# open data
DF <- fetch_survey(surveyID = 'HIDDEN',
                   force_request = T,
                   label = F,
                   convert = F)

# process data
DF <- DF %>%
  
  # remove incomplete or test observations
  filter(DistributionChannel != "preview",
         Progress > 95,
         !is.na(stim_lef_sma_mkn)) %>% 
  
  # select relevant variables
  select(
    # subject data
    ResponseId, sex, age, attention_check,
    
    # lab data
    lab, modality, incentive,
    
    # task data
    orientation, info,
    stim_lef_sma_mkn : prq_w_2) %>% 
  
  # identify conditions
  pivot_longer(cols = stim_lef_sma_mkn : stim_lef_sma_mky,
               names_to = "trial") %>% 
  rowwise() %>% 
  mutate(info.pres = 
           if_else(condition = grepl("mky", trial),
                   true = "yes",
                   false = "no"),
         just = 
           if_else(condition = grepl("lef", trial),
                   true = "left",
                   false = "right")) %>% 
  ungroup() %>% 
  
  # fix variable classes
  mutate_at(vars(info.pres, just),
            as.factor) %>% 
  
  # remove unnecessary rows
  filter(!is.na(value)) %>% 
  
  # remove incomplete participant
  filter(ResponseId != "R_1oI7B3iN2wVwaaj")
  
  # export
DF %>%
  write.csv("coles2024.csv",
            row.names = F)
```
